---
title: "Weighlighting Notebook"
author: "Nathan Chaney"
date: "2/11/2017"
output: html_document
---

# To Do

## Output
### Aggregate planned cycles for each lift to printable worksheet/schedule format

## Abstract to program/cycle level 
### number of cycles/programs dictates order of polynomial fitting
### compare error of polynomial fitting to ln fitting

# Narrative

I started lifting weights in March 2016 because my belly was too much larger than my shoulders, and I don't want to look like a frail old guy. I've lifted for a few months at a time before, but never stuck with it for more that 3 or 4 months. This time, however, I have been more diligent -- so much so, in fact, that I had to switch from a beginner program [(Stronglifts 5x5)](http://stronglifts.com/5x5/) to an intermediate schedule [(Madcow)](http://stronglifts.com/madcow-5x5-training-programs/) because I had stopped improving.

In Stronglifts 5x5, you alternate between 2 groups of 3 lifts (A: squat, bench, barbell row; B: squat, overtail press, and deadlift). You warm up, then lift 5 sets of 5 reps. When you successfully complete the lift, you add weight (2.5 lbs or 5 lbs, depending on the lift) the next time you perform the lift. If you fail, you deload. If you fail again, you reduce the number of sets and reps (5x3, 3x3, etc.).

Stronglifts is good for beginners like me because the weight starts low and you see rapid improvement. However, once the weight gets heavy (e.g., squatting 300 lbs), the volume of work simply gets to be too much for adequate recovery.

Madcow is a 12-week program similar to Stronglifts in that the work set (the max weight you use on a particular lift) is a heavy weight, low rep set. However, it is just a single set, and you perform incremental lifts to reach that set.

The Stronglifts app had charts that showed progress over time. However, I haven't found an app I like that lets you customize the lifts you're doing. I've seen things like Strongr that are customizable, but the learning curve seems steep and I'd rather apply that learning to R. 

## Libraries

```{r Setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
my.datadir <- "~/.bin/rweightlifting/raw-data/"
```

```{r Libraries, echo=FALSE, include=TRUE, warning=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(data.table)
library(plotly)
library(rweightlifting)

```

## Data Import & Munging

### Weightlifting Data

The load_csv_data() function reads all CSV files in the specified directory. Each file should correspond to a specific weightlifting program. The function expects a specific input format that includes a header row.

Future plans include the ability to import CSV data from the StrongLifts 5x5 app.

```{r}
# Expected CSV format:
# date,exercise,variant,set1weight,set1reps,set2weight,set2reps,set...weight,set...reps,setNweight,setNreps
# E.g., 2017-01-26,bench,barbell,102.5,5,135,5,155,5,180,5,210,3,155,13
data <- load_csv_data(datadir = my.datadir)
```

# Munge and Summarize

```{r include=TRUE, echo=TRUE}
start.date <- min(data$date, na.rm=TRUE)

exercises <- c("squat", "bench", "row", "tricep extension", "dips", "body weight", "deadlift", "press", "pullup", "bicep curl")

Comp.opt <- function(data)
{
    output <- integer(length(data)-1L)
    for(i in seq_along(output))
    {
        output[[i]] <- (data[[i]] >= 0.85 * data[[i+1L]])
    }
    return(output)
}

summary <- ddply(data, date ~ exercise, function(x)x[which.max(x$weight), ])
summary <- summary[with(summary, order(exercise, date)),]
true.max <- as.logical(Comp.opt(summary$weight))
summary <- subset(summary, true.max)

data$exercisef <- factor(data$exercise, levels = exercises, ordered=TRUE) 
summary$exercisef <- factor(summary$exercise, levels = exercises, ordered=TRUE) 

tail(summary)
tail(data)

```

# Tonnage

```{r}
library(dplyr)
tonnes <- data %>%
  filter(! exercisef %in% c("tricep extension", "bicep curl", "pullup", "dips")) %>%
  mutate(tonnage = weight * reps, date = as.POSIXct(paste(date, "18:00:00", sep=""), format = "%Y-%m-%d", tz = "America/Chicago"))

tonnage <- ggplot(tonnes) +
  geom_bar(aes(date, tonnage, fill = exercisef), stat = "identity") +
  scale_fill_brewer("Exercise", palette="Paired")

ggplotly(tonnage)

#NROW(unique(data$date))
```

# Summary by Program

```{r}
stronglifts.stop <- as.Date("2016-09-01")
stronglifts.tonnage <- data[data$date <= stronglifts.stop, ] %>%
  mutate(tonnage = weight * reps, date = as.POSIXct(paste(date, "18:00:00", sep=""), format = "%Y-%m-%d", tz = "America/Chicago")) %>%
  group_by(exercisef) %>% summarize(stronglifts.max = max(tonnage)) 

madcow.tonnage <- data[data$date > stronglifts.stop, ] %>%
  mutate(tonnage = weight * reps, date = as.POSIXct(paste(date, "18:00:00", sep=""), format = "%Y-%m-%d", tz = "America/Chicago")) %>%
  group_by(exercisef) %>% 
  summarize(madcow.max = max(tonnage))

tonnages.maxes <- merge(stronglifts.tonnage, madcow.tonnage, all = TRUE)
mdata <- melt(tonnages.maxes, id=c("exercisef"))
tonnages.maxes$diff <- tonnages.maxes$madcow.max - tonnages.maxes$stronglifts.max
tonnages.maxes$diff.percent <- round(tonnages.maxes$diff / tonnages.maxes$stronglifts.max * 100, 1)
# This will change to in SELECTED instead of ! IN this static list
tonnages.maxes <- tonnages.maxes %>%
  filter(! exercisef %in% c("tricep extension", "bicep curl", "pullup", "dips"))

ggplot(mdata) +
  scale_fill_brewer("Program", palette="Paired") +
  geom_bar(aes(exercisef, value, fill = variable, group=variable), stat = "identity", position = "dodge") +
  xlab("") +
  ylab("Max Tonnage (lbs)") +
  coord_flip()

ggplot(tonnages.maxes) +
  scale_fill_brewer("Exercise", palette="Paired") +
  geom_bar(aes(exercisef, diff.percent, fill = exercisef), stat = "identity", position = "dodge", na.rm = TRUE) +
#   theme(
# 	  axis.text.x = element_text(angle = 45),
#   ) +
  labs(title = "% Improvement (tonnage)", 
       subtitle = "Madcow over Stronglifts",
       x = "",
       y = "") +
  coord_flip() +
  theme(
    legend.position = "none"
  )

```

```{r}
#summary(data[data$date <= stronglifts.stop, ], by=exercisef)
stronglifts.max <- data[data$date <= stronglifts.stop, ] %>% group_by(exercisef) %>% summarize(stronglifts.max = max(weight))
madcow.max <- data[data$date > stronglifts.stop, ] %>% group_by(exercisef) %>% summarize(madcow.max = max(weight))

maxes <- merge(stronglifts.max, madcow.max, all = TRUE)
mdata <- melt(maxes, id=c("exercisef"))
maxes$diff <- maxes$madcow.max - maxes$stronglifts.max
maxes$diff.percent <- round(maxes$diff / maxes$stronglifts.max * 100, 1)

maxes <- maxes %>%
  filter(! exercisef %in% c("tricep extension", "bicep curl", "pullup", "dips"))

ggplot(mdata) +
  scale_fill_brewer(palette="Paired") +
  geom_bar(aes(exercisef, value, fill = variable, group=variable), stat = "identity", position = "dodge") +
  xlab("Exercise") +
  ylab("Max Weight (lbs)") +
  coord_flip()
  
ggplot(maxes) +
  scale_fill_brewer("Exercise", palette="Paired") +
  geom_bar(aes(exercisef, diff.percent, fill = exercisef), stat = "identity", position = "dodge", na.rm = TRUE) +
#   theme(
# 	  axis.text.x = element_text(angle = 45),
#   ) +
  labs(title = "% Improvement (max work set)", 
       subtitle = "Madcow over Stronglifts",
       x = "",
       y = "") +
  coord_flip() +
  theme(
    legend.position = "none"
  )
```

Need to calculate by tonnage

# Plotting

## Same size facets
## Could loop through each and do each exercise separately on same scale for carousel display

```{r}
today <- Sys.Date()
date.window <- seq(Sys.Date(), length = 2, by = "-6 months")[2]

plot.full <- ggplot(data=data, aes(date, weight, color=exercise, size=reps)) +
  scale_color_brewer(palette="Paired") +
  facet_wrap( ~ exercisef, ncol=5) +
  geom_point(data=data, aes(date, weight, color=exercise), alpha=0.5) +
  scale_radius(range = c(0.5, 4), breaks=c(3,6,9,12,15)) +
  scale_y_continuous(
    position="right",
    minor_breaks=function(x){
      seq(0, ceiling(max(x)), 25)
      }) +
  scale_x_date(
    date_breaks = "3 month", date_labels = "%b %y",
    date_minor_breaks = "1 month"
  ) +
  #coord_cartesian(xlim = c(date.window, today)) +
  coord_cartesian(xlim = c(start.date, today)) +
  geom_vline(aes(xintercept = as.numeric(Sys.Date())), color=alpha("green", 0.5), linetype="dashed")

plot.full.trendline <- plot.full +
  geom_line(data=summary, aes(date, weight, color=exercise), size=1, alpha=1) +
  labs(caption = paste("image generated at ", Sys.time(), sep = ""))

ggplotly(plot.full)
```
